### 2021-07-28 psy_peds.Rmd
first version of the R code to run the psy peds trend analysis

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra")
paket(listOfPackages)
```

# R session information
```{r}
sessionInfo()
```

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Inclusion criteria
The inclusion criteria for this study is:
- Patients inclusion criteria: 
  - Age between 11-17 at the date of data collection (*currently, subject to change in a future version)
- Visits inclusion criteria:
  - Inpatients 
  - Starting date between January 1st, 2019 and July 31st, 2021
  - Comprising least one ICD-10 code pertaining to the specified ICD-10 codes list (mental health related codes)

# Variables that need to be checked/modified by each site 
Change the values of the following variables according to the specificities of your site:
- "folder_4ce_files": folder path where your phase 2.2 data files are located
- "obfuscation": determine the obfuscation threshold (FALSE if no obfuscation, or the numeric value of the obfuscation threshold if any)
- "raceAvailable": set as TRUE or FALSE depending on whether the variable is being collected at your site
- "dateFormat": specify the format of the date at your site (e.g., for "03-AUG-20", the format would be "%d-%b-%y", [see documentation](https://www.stat.berkeley.edu/~s133/dates.html))

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/4ceData/Input"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
```

# Data input
We will use as input the 2.2 data files. Specifically: 
- LocalPatientSummary
- LocalPatientObservation
- LocalPatientClinicalcourse

For sites recording the race, we will also use an additional file:
- LocalPatientRace

## Read csv files

### 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

files <- readInputFiles( path      = folder_4ce_files, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

## Create the output folder if it doesn't exist
if (! "output" %in% list.dirs()) dir.create("output")

### Extract the patient summary and observation information. 
demo_raw <- files[[1]]
obs_raw <- files[[2]]
clinical_raw <- files[[3]]

### Read the file containing race information for those sites recording this variable
if( raceAvailable == TRUE ){
  race_raw <- read.delim(paste0( folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
}
```

### Read the ICD10 codes file
The file containing the ICD10 psychiatric related codes is located in the `public-data` folder of the GitHub repository.

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
  rename(disorder_group = Mental.Health.Disorder.Group, 
         description = Description, 
         originalCode = ICD10_Code) %>%
  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
  select( disorder_group, ICD10_Code, description)
```

# Data management
We will select only patients that have been diagnosed with at least one of the mental health ICD10 codes at any point during study followup. 
For these patients we will generate a table that will allow us to perform all the analysis required for this study. 
This transformed table will contain 9 columns:
- patient_num: patient internal identifier
- sex (male, female)
- age (in years)
- concept_type (DIAG-ICD10)
- concept_code (ICD-10 full code, e.g, R44.1)
- calendar_date (day-month-year, 30-SEP-20)
- hospitalization status (1, 0)
- n_hospitalization (the hospitalization number)
- len_hospitalization (length of each hospitalization)

The last two columns are not available in the input data files. We will generate those counts using the information available in the clinical raw file.

## Adding columns with hospitalization number and hospitalization length to the clinical raw file
```{r}
count_sequences_hospitalisation <- function(df, ...) {
  count_sequences <- rle(df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  df[df$in_hospital == 1, "n_hospitalisation"] <- sequences
  df[df$in_hospital == 1, "len_hospitalisation"] <- sequences_len
  return(df)
}
clinical_raw <- clinical_raw %>%
  group_by(patient_num) %>% 
  group_modify(count_sequences_hospitalisation)
```

## Creating the inputTable
Creation of the inputTable that will be the only table used throughout the rest of the analysis
- selecting only visits with at least one ICD codes pertaining to the mental health categories
- filter visits where patients are between 11 and 17 years old
- add the demographic and hospitalization related information 
- add the ICD code description and the related disorder group

We will also transform the observation dates, and break it by weeks, adding 2 additional columns: "weeks" and "year". Finally, we will dichotomize the study timeframe in two periods, based on cut-off dates, "before_pandemic", and "after_pandemic".

Additionally, we will also add the information about race to the table for sites collecting this information.

### Check that we have patients with the ICD10 mental codes
```{r}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
print(codesToReview)

### Estimate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( obs_raw %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 concept_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( concept_code = as.character( concept_code ), 
                 concept_length = nchar( concept_code ), 
                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( concept_code, concept_length, concept_check) )

summary(as.factor( mentalHealthCodes$concept_check ) )

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Generating the table
```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
site <- unique(as.character(demo_raw$siteid))

# cut-off dates used in the analysis 
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
#early_pandemic_start_date <- "2020-03-15"
#late_pandemic_start_date <- "2021-01-01"

input_data <- obs_raw %>% 
  full_join( clinical_raw,  by = c("patient_num", "days_since_admission")) %>%
  full_join( demo_raw, by = "patient_num") %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = cut( date, breaks= "week"),
                 year = format( date, "%Y")) %>% 
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 in_hospital == 1,
                 concept_code %in% icdCodes$ICD10_Code,
                 age < 18 & age > 10,
                 date >= history_start_date & date <= history_end_date) %>%
  # dplyr::mutate(period = ifelse( date < early_pandemic_start_date,
  #                                 "before_pandemic",
  #                                 ifelse(date < late_pandemic_start_date,
  #                                        "early_pandemic", 
  #                                        "late_pandemic"))) %>%
  dplyr::mutate(period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","after_pandemic")) %>%
  left_join( icdCodes, by = c("concept_code" = "ICD10_Code")) %>%
  dplyr::select( patient_num, sex, age, concept_type, concept_code, calendar_date, in_hospital, n_hospitalisation, len_hospitalisation, 
                 date, weeks, year, period, disorder_group, description )
if( raceAvailable == TRUE ){
  input_data <- input_data %>%
    left_join( race_raw, by = "patient_num" )
}
```

# Data analysis

## Patients having presented history of mental health related issues before pandemic
We estimate the number of patients that presented a mental health disorder (grouped by mental health disorder groups - icdCodes$disorder_group):
- only before the pandemic
- only after the pandemic
- both before and after pandemic

The output is a barplot showing in the x-axis the number of patients, and in the y-axis the distinct disorder groups. Each disorder group has 3 bars associated based on the three previous options.

```{r}
patient_count_barplot_period <- input_data %>% 
  dplyr::group_by( disorder_group, patient_num ) %>% 
  dplyr::mutate( 
    when = ifelse( "before_pandemic" %in% period & "after_pandemic" %in% period,
                   "Before and After Pandemic",
                   ifelse( "before_pandemic" %in% period & ! "after_pandemic" %in% period,
                           "Only Before Pandemic",
                           "Only After Pandemic"
                   )
    ) 
    %>% factor(levels = c("Only Before Pandemic", "Before and After Pandemic", "Only After Pandemic"))
  ) %>%
  dplyr::ungroup() %>% 
  dplyr::select( patient_num, disorder_group, when ) %>%
  dplyr::group_by( disorder_group, when ) %>%
  dplyr::summarise( count = n_distinct( patient_num ) ) %>% 
  dplyr::mutate(count = ifelse(count > obfuscation | isFALSE(obfuscation),
                               count, 
                               0.5), 
                siteid = site)

patient_count_barplot_period %>%
  ggplot(aes(x = as.factor( disorder_group), y = count, fill = when)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_manual(values = cbPalette, name = "Period", guide = guide_legend(reverse = TRUE)) +
  labs(y = "Unique patients count",
       x = "Disorder group",
       title = "Patients with at least one visit", 
       subtitle = "According to visit period and pathology group", 
       legend = "Period")+
  theme(legend.position = "bottom", 
        legend.direction = "vertical",
        title = element_text(size=10),
        plot.subtitle = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size =7, angle = 0, hjust = 0.5, vjust = 0)) + 
  coord_flip()

ggsave("output/disorder_before_after_pandemic_all.png", width = 18, height = 9, units="cm")
```

## Diagnoses count per week
Here we calculate the number of patients hospitalized per week (exactly, the number of patients being assigned a diagnosis on that week, hence usually corresponding to the end of hospitalization, ie not taking into account the length of the encounter). 
The barplot shows the weeks on the x-axis, and the patient counts on the y-axis. The horizontal lines is the (quadratic) temporal trend of these patient counts.
```{r}
patient_count_per_week <- input_data[, c("patient_num", "weeks")] %>%
  unique() %>%
  group_by(weeks) %>%
  tally() %>%
  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
                     n,
                     0.5),
         siteid = site )

#estimate the number of patients per week
patient_count_per_week %>%
  dplyr::mutate(weeks = as.Date(weeks)) %>% 
  dplyr::mutate(period = ifelse( weeks < pandemic_start_date,
                                  "before_pandemic",
                                  #ifelse(weeks < late_pandemic_start_date,
                                  #       "early_pandemic", 
                                  #       "late_pandemic")
                                 "after_pandemic")) %>% 
  ggplot(aes(x = weeks, y = n, fill = period)) +
  geom_bar(stat="identity")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Calendar Weeks",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")
```

## Temporal trend of hospitalized patients per day

```{r, }
# Creating encounter table with observation assigned to encounters
# 1. Creating encounters identifiers based on clinical_raw information about start and end date of an encounter --> n_hospitalisation
# 2. Assigning ICD codes to encounters (because they are assigned to the first or last day but usually are meaningful for the entire encounter)
encounters_obs <- clinical_raw %>% 
  filter(in_hospital == 1) %>% 
  select(patient_num, days_since_admission, n_hospitalisation) %>% 
  left_join(obs_raw[obs_raw$concept_type == "DIAG-ICD10",], on = c("patient_num", "days_since_admission")) %>% 
  drop_na(concept_type) %>% 
  distinct(patient_num, n_hospitalisation, concept_type, concept_code)

obs_by_day <- encounters_obs %>% 
  filter(concept_code %in% icdCodes$ICD10_Code) %>% 
  left_join(clinical_raw[clinical_raw$in_hospital == 1, c("patient_num", "n_hospitalisation", "calendar_date")], 
            on = c("patient_num", "n_hospitalisation")) %>% 
  group_by(calendar_date) %>% 
  summarise(patient_count = n_distinct(patient_num)) %>% 
  mutate(calendar_date = as.Date(calendar_date, format=dateFormat), 
         period = if_else(calendar_date < pandemic_start_date,
                          "Before Pandemic",
                          #if_else(calendar_date < late_pandemic_start_date,
                          #        "Early Pandemic", 
                          #        "Late Pandemic")
                          "After Pandemic"
         ), 
         patient_count = ifelse(patient_count > obfuscation | isFALSE(obfuscation), 
                                 patient_count, 
                                 0.5)
  )

ggplot(obs_by_day, aes(x = calendar_date,
                       y = patient_count,
                       color = period))  +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  #geom_vline(xintercept = as.Date(late_pandemic_start_date),
  #           linetype = "dashed") +
  scale_color_manual(values = cbPalette) +
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Daily Hospitalized Patient Counts", 
       subtitle = "Hospitalisation encounters with at least one mental health diagnosis",
       legend = "Period")
ggsave("output/hospitalized_patients_all.png", width = 18, height = 9, units="cm")
```

### Patients count per week aggregated by sex
Same graphics as "Patient Count per Week", dichotomized by patients sex.
```{r}
patient_count_per_week_sex <- input_data[, c("patient_num", "sex","weeks")] %>%
  unique() %>%
  group_by(weeks, sex) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation),
                     n, 
                     0.5),
         siteid = site)

patient_count_per_week_sex %>%
  ggplot(aes(x = as.Date(weeks), y = n, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Week",
       title = "Weekly patient counts with mental health related ICD codes aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

### Patients count per week aggregated by age group 
This graphic will only make sense if we include a wide range of age.

```{r}
#estimate the number of patients per week by age group
input_data[, c("patient_num", "age","weeks")] %>%
  unique() %>%
  mutate( age_group = ifelse( age < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
  group_by(weeks, age_group) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation, n, obfuscation * 0.5))  %>%
    ggplot(aes(x = as.Date(weeks), y = n, fill = age_group)) +
    geom_bar(stat="identity", position = "dodge")+
    labs(y = "Counts",
         x = "Week",
         title = "Weekly patient counts with mental health related ICD codes aggregated by age group",
         subtitle = "Per age group") + 
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = cbPalette) +
    geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

```

## Patients counts per week grouping by ICD subgroup category 
Here we estimate the number of patients hospitalized per week but we break it based on the distinct disorder groups. 
The output is a barplot showing in the x-axis the week, and in the y-axis the patient counts. Additionally a temporal trend line is added on top of the bars.
Each barplot represent a disorder group.

```{r}

patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1,c("patient_num", "weeks", "disorder_group")] %>%
  unique() %>%
  group_by(weeks, disorder_group) %>% tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation), n, 0.5), siteid = site)

patient_count_per_week_disorder_group %>%
  mutate(weeks = as.Date(weeks), 
         period = if_else(weeks < pandemic_start_date,
                          "Before Pandemic",
                          #if_else(weeks < late_pandemic_start_date,
                          #        "Early Pandemic", 
                          #        "Late Pandemic")
                          "After Pandemic"
         )
  ) %>% 
  ggplot(aes(x = as.Date(weeks), y= n, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  #geom_vline(xintercept = as.Date(late_pandemic_start_date),
  #           linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Week",
       y = "Patient Counts",
       title = "Weekly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
ggsave("output/categorical_counts_inpatient_teens_all.png", width = 18, height = 9, units="cm")
```

# Summary table 1
Here we will estimate the counts to populate table 1, with the counts for early and late pandemic, and a statistical test comparing both. 
This table one will include:
- sex
- age
- race (when available)
- hospitalization length
- diagnosis codes 
- diagnosis groups

```{r}
t.test2 <- function(m1, m2, s1, s2, n1, n2) {
  # t.test computed from sample statistics, ie using mean, standard deviation and count value of the samples
  # useful in case we can only retrieve aggregated stats from the sites
  # ref: https://stats.stackexchange.com/questions/30394/how-to-perform-two-sample-t-tests-in-r-by-inputting-sample-statistics-rather-tha 
  # and https://en.wikipedia.org/wiki/Student%27s_t-test?oldformat=true#Equal_or_unequal_sample_sizes,_unequal_variances_(sX1_%3E_2sX2_or_sX2_%3E_2sX1)
  se <- sqrt( (s1^2/n1) + (s2^2/n2) )
  # welch-satterthwaite df
  df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
  t <- (m1-m2)/se 
  dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))    
  names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
  return(dat)
}
```

## Age
```{r}
age_stats <- input_data %>%
  dplyr::distinct(patient_num, period, age) %>%
  filter(age != 0) %>% 
  dplyr::group_by( period ) %>% 
  dplyr::summarise( mean = mean( age ), 
                    sd = sd( age ), 
                    count = n()) %>% 
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::ungroup() %>% 
  dplyr::bind_rows(
    dplyr::summarise(input_data, period="Total", mean=mean(age), sd=sd(age), count = n())
  )

age_test <- t.test2(m1 = age_stats[[1, "mean"]], 
                    m2 = age_stats[[2, "mean"]], 
                    s1 = age_stats[[1, "sd"]], 
                    s2 = age_stats[[2, "sd"]], 
                    n1 = age_stats[[1, "count"]], 
                    n2 = age_stats[[2, "count"]]) %>% 
  `[[`("p-value")
```

## Sex
```{r}
sex <- input_data %>% 
  dplyr::group_by( sex ) %>%
  dplyr::summarise( count = n_distinct(patient_num) ) %>%
  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                 total = sum( count ), percentage = round( count/total*100, 2)  )

sex_by_period <- input_data %>% 
  group_by(period, sex ) %>% 
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, .add=TRUE) %>% 
  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
          percentage=round(count/sum( count )*100, 2)) %>% 
  replace_na(list(male = 0, female = 0))

sex_test <- sex_by_period %>% 
  select(-percentage) %>% 
  pivot_wider(names_from = sex, values_from = count) %>% 
  ungroup() %>% 
  select(male, female) %>% 
  chisq.test() %>% 
  `[[`("p.value")
```

## Race
```{r}
if( raceAvailable == TRUE ){
  race <- input_data %>%
    dplyr::group_by( race_4ce ) %>%
    dplyr::summarise( count = n_distinct(patient_num) ) %>%
    dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                   total = sum( count ), percentage = round( count/total*100, 2)  )

  race_by_period <- input_data %>%
    group_by(period, race_4ce ) %>% 
    summarise(count = n_distinct(patient_num)) %>%
    group_by(period, .add=TRUE) %>% 
    mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
            percentage=round(count/sum( count )*100, 2))

summary_race_test <- race_by_period %>% 
  select(-percentage) %>% 
  tidyr::pivot_wider(names_from = period,
                     values_from = count) %>%
  select(after_pandemic) %>% 
  tidyr::replace_na(list(after_pandemic = 0)) %>% 
  chisq.test() %>% 
  `[[`("p.value")
}
```

## Diagnosis

### By ICD code
```{r}
summary_diagnosis <- input_data %>%
  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>% 
  dplyr::group_by( concept_code, description, period ) %>%
  dplyr::count( patient_num ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc(count) ) %>% 
  dplyr::ungroup()

summary_diagnosis_test <- summary_diagnosis %>% 
  tidyr::pivot_wider(names_from = period,
                     values_from = count) %>%
  #select(late_pandemic, early_pandemic) %>% 
  select(after_pandemic) %>%
  #tidyr::replace_na(list(late_pandemic = 0, early_pandemic = 0)) %>% 
  tidyr::replace_na(list(after_pandemic = 0)) %>% 
  chisq.test() %>% 
  `[[`("p.value")

```

### By disorder group
```{r}
summary_disorder_group <- input_data %>%
  dplyr::group_by( disorder_group ) %>%
  dplyr::count( patient_num ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc( count ))

summary_disorder_group_by_period <- input_data %>% 
  dplyr::distinct( patient_num, disorder_group, period, weeks ) %>% 
  dplyr::group_by( period, disorder_group ) %>%
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, add=TRUE) %>%
  mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
         percentage=round(count/sum( count )*100, 2))

summary_disorder_group_by_period_test <- summary_disorder_group_by_period %>% 
  pivot_wider(id_cols = disorder_group, names_from = period, values_from = c(count, percentage)) %>% 
  mutate(across(where(is.numeric), ~replace_na(.x, replace = 0))) %>% 
  #select(count_early_pandemic, count_late_pandemic) %>% 
  select(count_after_pandemic) %>%
  chisq.test() %>% 
  `[[`("p.value")

```

## Hospitalization

### Length of hospitalization 
```{r}
## Calculating also mean and sd in case we can only retrieve aggregated data from the sites
hospit_length <- input_data %>%
  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
  dplyr::group_by( period ) %>% 
  dplyr::summarise( n_hospitalisations = n(),
                    length_hospt_median = median(len_hospitalisation), 
                    IQR = IQR( len_hospitalisation), 
                    length_hospt_mean = round(mean(len_hospitalisation), 2),
                    length_hospt_sd = round(sd(len_hospitalisation), 2))

hospit_length_mean_test <- t.test2(
  m1 = hospit_length[[1, "length_hospt_mean"]],
  m2 = hospit_length[[2, "length_hospt_mean"]],
  s1 = hospit_length[[1, "length_hospt_sd"]],
  s2 = hospit_length[[2, "length_hospt_sd"]],
  n1 = hospit_length[[1, "n_hospitalisations"]],
  n2 = hospit_length[[2, "n_hospitalisations"]]
)

```

### Repeat hospitalizations by patient
```{r}
hospitalisation_count <- input_data %>% 
  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
  dplyr::group_by( period, patient_num ) %>%
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>%
  dplyr::group_by( period, add = FALSE) %>% 
  dplyr::summarise( mean_count_hospitalizations = mean(n_hospitalisation), 
                    max_count_hospitalizations = max(n_hospitalisation),
                    median_count_hospitalizations = median(n_hospitalisation), 
                    sd_hospitalisation_count = sd(n_hospitalisation),
                    IQR_hospitalisation_count = IQR( n_hospitalisation ))

repeated_hospitalisation <- input_data %>% 
  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
  dplyr::group_by(patient_num, period) %>% 
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>% 
  dplyr::mutate(repeated_hospitalisation = if_else(n_hospitalisation > 1, "repeated", "non_repeated")) %>% 
  dplyr::group_by( period, add = FALSE ) %>% 
  dplyr::count( repeated_hospitalisation ) %>% 
  tidyr::pivot_wider(names_from = repeated_hospitalisation, values_from = n) %>% 
  replace_na(list("non_repeated" = 0, "repeated" = 0))
repeated_hospitalisation_test <- chisq.test(repeated_hospitalisation[, c("non_repeated", "repeated")]) %>% 
  `[[`("p.value")
```

## Pulling together Table 1 statistics
```{r}
format_counts <- function(count, percentage) {
  glue::glue("{count} ({percentage} %)", count = count, percentage = percentage)
}

sex_table1 <- sex_by_period %>% 
  bind_rows(
    sex_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(sex = "Total", count = sum(count), percentage = sum(percentage))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  rename(name = sex)

if( raceAvailable == TRUE ){
  race_table1 <- race_by_period %>% 
  bind_rows(
    race_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(race_4ce = "Total", count = sum(count), percentage = sum(percentage))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  rename(name = race_4ce)
}

diagnosis_table1 <- summary_diagnosis %>% 
    bind_rows(
    sex_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(concept_code = "Total", count = sum(count))
  ) %>%
  pivot_wider( names_from = period, values_from = count) %>% 
  mutate(diag = paste(concept_code, description, sep = ": ")) %>% 
  select(diag, after_pandemic) %>% 
  rename(name = diag)

disorder_table1 <- summary_disorder_group_by_period %>% 
  bind_rows(
    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  # replace_na( list(early_pandemic = "0 (0 %)", 
  #                  late_pandemic = "0 (0 %)")) %>% 
    replace_na( list(after_pandemic = "0 (0 %)")) %>% 
  rename(name = disorder_group)

hospit_table1 <- hospit_length %>% 
  mutate(median = glue::glue("{median} ({IQR})", median = length_hospt_median, IQR = IQR), 
         mean = glue::glue("{mean} ({sd})", mean = length_hospt_mean, sd = length_hospt_sd), 
         n_hospitalisations = as.character(n_hospitalisations)) %>% 
  select(median, mean, n_hospitalisations, period) %>% 
  pivot_longer(cols = c("n_hospitalisations", "median", "mean")) %>% 
  pivot_wider(names_from = period, values_from = value)

repeated_hospit_table1 <- repeated_hospitalisation %>% 
  pivot_longer(cols = c("non_repeated", "repeated")) %>% 
  pivot_wider(names_from = period, values_from = value)

list_tests <- c("Sex" = sex_test, 
             "Hospitalisations length" = repeated_hospitalisation_test, 
             "Repeated hospitalisations" = repeated_hospitalisation_test,
             "Diagnoses" = summary_diagnosis_test, 
             "Disorder groups" = summary_disorder_group_by_period_test)

list_table1 <- list(
    "Sex" = sex_table1, 
    "Hospitalisations length" = hospit_table1,
    "Repeated hospitalisations" = repeated_hospit_table1,
    "Disorder groups" = disorder_table1, 
    "Diagnoses" = diagnosis_table1)
if (raceAvailable) {
  list_tests["Race"] = summary_race_test
  list_table1 = c(list("Race" = race_table1), 
          list_table1
  )
}

tests <- list_tests %>% 
  data.frame() %>% 
  round(2)
names(tests) <- "pvalue"
tests["Category"] <- row.names(tests)

table1 <- list_table1 %>% 
  lapply(function(df) mutate(df, across(.fns = as.character))) %>% 
  dplyr::bind_rows(.id = "Category") %>% 
  dplyr::left_join(tests, by = "Category")

table1 %>% 
  group_by(Category) %>%
  mutate(pvalue = ifelse(duplicated(pvalue, fromLast = TRUE), "", as.character(pvalue))) %>%
  ungroup() %>%
  mutate(Category = "") %>%
  mutate(across(.cols = c(before_pandemic, after_pandemic), .fns = replace_na, "0")) %>% 
  rename_with(str_to_title) %>%
  rename_with(sub, pattern = "_", replacement = " ") %>%
  mutate(across(.fns = str_to_title)) %>% 
  mutate(across(.fns = sub, pattern = "_", replacement = " ")) %>% 
  kable( booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
```

# Save the output in a RData file. 
```{r}

length_hospitalisation_values <- input_data %>% 
  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation ) %>%
  dplyr::mutate( siteid = site ) %>%
  dplyr::select( len_hospitalisation, siteid )

save( patient_count_barplot_period,
      patient_count_per_week,
      patient_count_per_week_sex,
      patient_count_per_week_disorder_group,
      table1,
      length_hospitalisation_values,
      obs_by_day,
      file = paste0("output/peds_psy_aggResults_", unique(as.character(demo_raw$siteid)), ".RData"))
```