### 2021-10-01 psy_peds_3.Rmd
obfuscation added and time period as a variable

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
paket(listOfPackages)
```

# R session information

```{r}
sessionInfo()
```

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Inclusion criteria

The inclusion criteria for this study is:

-   Patients inclusion criteria:

    -   Age between 11-17 at the date of visit

-   Visits inclusion criteria:

    -   Inpatients (at least one hospitalization \> 1 day)

    -   Start date: January 1st, 2019

    -   End date: May 31st, 2021

    -   Comprising at least one ICD-10 code during hospitalization

# Variables that need to be checked/modified by each site

Change the values of the following variables according to the
specificities of your site:

1.  "folder_4ce_files": folder path where your phase 2.2 data files are
    located

2.   "obfuscation": determine the obfuscation threshold (FALSE if no
    obfuscation, or the numeric value of the obfuscation threshold if
    any)

3.   "raceAvailable": set as TRUE or FALSE depending on whether the
    variable is being collected at your site

4.   "dateFormat": specify the format of the date at your site (e.g., for
    "03-AUG-20", the format would be "%d-%b-%y", [see
    documentation](https://www.stat.berkeley.edu/~s133/dates.html))

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
time_period <-"weeks" #possible values: days, weeks, months
```

```{r}
# cut-off dates used in the analysis, these should not be changed
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
start_date_plots <- "2019-01-15"
end_date_plots <- "2021-05-01"
```

# Data input

We will use as input the 2.2 data files. Specifically:

- LocalPatientSummary

- LocalPatientObservation

- LocalPatientClinicalcourse

For sites recording the race, we will also use an additional file:

- LocalPatientRace

## Read csv files

### 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

files <- readInputFiles( path      = folder_4ce_files, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

## Create the output folder if it doesn't exist
#if (! "output" %in% list.dirs()) dir.create("output")

### Extract the patient summary and observation information. 
demo_raw <- files[["patientSummary"]] %>% filter(cohort == "AllAdm")
obs_raw <- files[["patientObservations"]] %>% filter(cohort == "AllAdm")
clinical_raw <- files[["patientClinicalCourse"]] %>% filter(cohort == "AllAdm")

### Read the file containing race information for those sites recording this variable
if( raceAvailable == TRUE ){
  race_raw <- read.delim(file.path(folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
}
```

### Read the ICD10 codes file

The file containing the ICD10 psychiatric related codes is located in
the `public-data` folder of the GitHub repository.

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   dplyr::select( disorder_group, ICD10_Code, description )
```

## Data-management

### Filtering on "inpatient" status

```{r}
# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
clinical_raw <- filter(clinical_raw, in_hospital == 1)
```

### Calculating duration of individual hospitalizations

Adding columns with hospitalization number and hospitalization length to the clinical raw file

```{r}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)

clinical_raw <- left_join(clinical_raw,
                          hospitalisations_seq_df,
                          by = c("patient_num", "cohort", "days_since_admission"))
```

### Counting ICD10 mental health codes in the obs_raw table

```{r}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
print(codesToReview)

### Calculate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( obs_raw %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 concept_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( concept_code = as.character( concept_code ), 
                 concept_length = nchar( concept_code ), 
                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( concept_code, concept_length, concept_check) )

summary(as.factor( mentalHealthCodes$concept_check ) )

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

## Applying inclusion criteria to filter and format the data

We are selecting hospitalizations based on the following criteria:

- Associated with at least 1 ICD code

- With a duration of hospitalization spanning over at least two consecutive days

- Starting on or after 2019/01/01

### Calculating age at hospitalization time
```{r}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))
stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
```

``` {r}
obs_filtered_age <- obs_raw_age %>% 
  filter(concept_type == 'DIAG-ICD10', 
         days_since_admission >= 0)

clinical_filtered <- clinical_raw %>%
  filter( len_hospitalisation > min_hosp_days, 
          cohort  %in% c("AllAdm")) %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic", "during_pandemic"))

clinical_filtered$time_p <- NA
if( time_period == "weeks"){
  clinical_filtered$time_p <- clinical_filtered$weeks
}else if(time_period == "days"){
  clinical_filtered$time_p <- clinical_filtered$date
}else if(time_period == "months"){
  clinical_filtered$time_p <- clinical_filtered$month
  }
```

```{r}
# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered_age %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
```

```{r}
stopifnot(unique(all_codes$cohort) == "AllAdm")
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
stopifnot(nrow(all_codes) == nrow(observation_table))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
```

## First estimation: total number of patients without filtering by any disease subtype

```{r}
admission_plot_count <- function(df,
                               x,
                               y, 
                               title,
                               y_title,
                               size = NULL, 
                               color = NULL){
  plot <- ggplot(df, aes_string(x = x, y = y, color = color, size = size))  +
    geom_point() +
    scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y", limits = c(start_date_plots, end_date_plots), expand = expansion(mult = 0.01, add = 1)) +
    theme(axis.text.x=element_text(angle=60, hjust=1))+
    labs(x = "Calendar date",
         y = y_title, 
       title = title) 
  return(plot)
}

```

### New hospitalizations

```{r}
new_hospitalizations <- clinical_table %>%
  group_by(patient_num, n_hospitalisation) %>%
  slice(which.min(time_p)) %>%
  filter( in_hospital == 1 ) %>%
    group_by( time_p ) %>%
  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>% 
  ungroup()

start_date_plots <- as.Date("2019-01-01")
end_date_plots <- as.Date("2021-05-24")

nh <- admission_plot_count(df = new_hospitalizations, 
                     x = "time_p",
                     y = "count", 
                     title = paste0("New admissions (per ", time_period,")"), 
                     y_title = "Patient Count")

nh
```

### Filtering on age at hospitalization time

Keeping only patients with age between 11 and 17

```{r}
# Filtering hospitalizations where patient age is in inclusion criteria
observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

# Filtering observation tables which patient_num and days_since_admission are still in clinical table after filtering on age

clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
            by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation, age_time_diagnosis)

clinical_table_ado <- clinical_table %>% 
  inner_join(clinical_test,
             by = c("patient_num", "n_hospitalisation"))

# Filtering observation tables which patient_num are still in clinical table after filtering on age
demographic_table_ado <- demographic_table %>% 
  inner_join(distinct(clinical_table_ado, patient_num), by = "patient_num")

all_codes_ado <- all_codes %>% 
   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
```

```{r}
stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
stopifnot(unique(all_codes_ado$in_hospital) == 1)
stopifnot(unique(all_codes_ado$len_hospitalisation > min_hosp_days))
stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
```

### New admissions age range 11-17

```{r}
new_hospitalizations_ado <- clinical_table_ado %>%
  group_by(patient_num, n_hospitalisation) %>%
  slice(which.min(time_p)) %>%
  filter( in_hospital == 1 ) %>%
    group_by( time_p ) %>%
  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>% 
  ungroup()

start_date_plots <- as.Date("2019-01-01")
end_date_plots <- as.Date("2021-05-24")

nh_ado <- admission_plot_count(df = new_hospitalizations_ado, 
                     x = "time_p",
                     y = "count", 
                     title = paste0("New admissions (per ", time_period, ")"), 
                     y_title = "Patient Count")+
    labs( subtitle = "(age range 11-17)")

nh_ado
```

### Total hospitalize patients age range 11-17

```{r}
unique_patients_hospitalized <- clinical_table_ado %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

th <- admission_plot_count(df = unique_patients_hospitalized, 
                     x = "time_p",
                     y = "count", 
                     title = paste0("Total hospitalization (per ", time_period, ")"), 
                     y_title = "Patient Count") +
  labs( subtitle = "(age range 11-17)")
th
```

## Total number of patients filtering by psychiatric disorders

## Creating the inputTable

Creation of the inputTable that will be the only table used throughout the rest of the analysis - selecting only visits with at least one ICD
codes pertaining to the mental health categories - filter visits where patients are between 11 and 17 years old - add the demographic and
hospitalization related information - add the ICD code description and the related disorder group

### Add the psychiatric/mental health ICD code description and filter

```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
site <- unique(as.character(demo_raw$siteid))

icdCategory <- read.delim("public-data/icd10Codes.txt",
                          header = FALSE,
                          colClasses = "character")
colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")

all_codes_desc <- all_codes_ado %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, time_p, disorder_group, description, icd_code_category )

all_codes_desc <- all_codes_desc %>% 
  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))

all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 
all_codes_psy <- all_codes_desc %>% 
  filter(concept_code %in% icdCodes$ICD10_Code)

if( raceAvailable == TRUE ){
  all_codes_psy <- all_codes_psy %>%
    left_join( race_raw, by = c("patient_num") )
}
```

### Total patients hospitalized comparing psychiatric disorders vs other

```{r}
count_icd_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count_psy =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd <- count_icd %>%
  left_join(count_icd_psy, 
            by = c("time_p", "period")) %>% 
  replace_na(list(count_psy = 0))

count_icd$percentage_psy <- 100*count_icd$count_psy / count_icd$count_icd

thpd_psyCount <- count_icd %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount

```

```{r}
thpd_psyPer <- count_icd %>%
  ggplot(aes(x = time_p, y = percentage_psy, fill=period)) +
  scale_fill_manual(values = cbPalette[c(3,4)]) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage of patients with mental health related ICD codes (per ", time_period,")")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
  geom_point()
thpd_psyPer
```

#### Psychiatric conditions aggregating by suptype

##### Total hospitalizations

```{r}
count_disorder_group <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p, disorder_group ) %>%
  summarise(count_dg =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg <- count_disorder_group %>%  
  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient Count",
       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg
```

Percentage using as denominator total number of patients

```{r}
# estimate the counts
count_icd <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group <- count_icd %>%
  left_join(count_disorder_group, 
            by = c("time_p", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group$percentage_dg <- 100*perc_disorder_group$count_dg / perc_disorder_group$count_icd

dgPerc <- perc_disorder_group %>%  
  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient (%)",
       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc
```

### Aggregated by sex

```{r}
# counts per time period and sex
patients_hospitalized_agg_sex <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( time_p, sex ) %>%
  summarise(count_sex =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

patients_hospitalized_agg_sex %>%
  ggplot(aes(x = time_p, y = count_sex, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = time_period,
       title = paste0("Patient counts per ", time_period, " aggregated by sex"),
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

##percentage
total_patients_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

patients_hospitalized_agg_sex_perc <- total_patients_psy %>%
  left_join( patients_hospitalized_agg_sex, by = c("time_p")) %>% 
  replace_na(list(count_sex = 0))

patients_hospitalized_agg_sex_perc$percentage <- 100*patients_hospitalized_agg_sex_perc$count_sex/patients_hospitalized_agg_sex_perc$count

patients_hospitalized_agg_sex_perc %>%
  ggplot(aes(x = time_p, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage ",
       x = time_period,
       title = paste0("Patient (%) per ", time_period," aggregated by sex"),
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```