### 2022-04-06 psy_peds_3_LURIE.Rmd

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Setting up the R environment

### Installing packages and loading the library

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}

#listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
#paket(listOfPackages)

# trying to remove kableExtra to run w/o docker
listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "tsModel", "gridExtra", "dplyr")
paket(listOfPackages)
```

### R session information

```{r}
sessionInfo()
```

### Set working directory

```{r}
#setwd("~/projects/nu4ce/4ceData_files/Phase2.2.psy_peds_trends")
```

### Plots aesthetics information

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Data retrieval

## Inclusion criteria
The inclusion criteria for this study is:

-   Patients inclusion criteria:

    -   Age between 11-17 at the date of visit

-   Visits inclusion criteria:

    -   Inpatients (at least one hospitalization \> 1 day)

    -   Start date: January 1st, 2019

    -   End date: May 31st, 2021

    -   Comprising at least one ICD-10 code during hospitalization

## Variables that need to be checked/modified by each site

Change the values of the following variables according to the
specificities of your site:

1.  "folder_4ce_files": folder path where your phase 2.2 data files are
    located

2.   "obfuscation": determine the obfuscation threshold (FALSE if no
    obfuscation, or the numeric value of the obfuscation threshold if
    any)

3.   "raceAvailable": set as TRUE or FALSE depending on whether the
    variable is being collected at your site

4.   "dateFormat": specify the format of the date at your site (e.g., for
    "03-AUG-20", the format would be "%d-%b-%y", [see
    documentation](https://www.stat.berkeley.edu/~s133/dates.html))

5. "data_update_date": date at which the data has been updated in the local 
datawarehouse. Used to estimate patient age at time of visit, since patients age
in the 4CE demographic file is expected the age at data update.

6. "time_period": Time unit for the different graphs of the plots: months (default), weeks, days

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "../4ce_ped_psych/data"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%m-%y" #"%d-%b-%y"
min_hosp_days <- 0
data_update_date <- "2021-06-05"
time_period <-"months" #possible values: days, weeks, months
```

```{r}
# cut-off dates used in the analysis, these should not be changed
#start_date_plots <- as.Date("2019-01-01")
#history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2018-12-01")
end_date_plots <- as.Date("2021-07-01")
```

## Data loading

We will use as input the 2.2 data files. Specifically:

- LocalPatientSummary

- LocalPatientObservation

- LocalPatientClinicalcourse

For sites recording the race, we will also use an additional file:

- LocalPatientRace

### Reading 4CE phase 2.2 files

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

files <- readInputFiles( path      = folder_4ce_files, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

## Create the output folder if it doesn't exist
#if (! "output" %in% list.dirs()) dir.create("output")

### Extract the patient summary and observation information. 
demo_raw <- files[["patientSummary"]] %>% filter(cohort == "AllAdm")
obs_raw <- files[["patientObservations"]] %>% filter(cohort == "AllAdm")
clinical_raw <- files[["patientClinicalCourse"]] %>% filter(cohort == "AllAdm")

### Read the file containing race information for those sites recording this variable
if( raceAvailable == TRUE ){
  race_raw <- read.delim(file.path(folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
}
```

### Read in Lurie files

```{r}
non_psych_inpt_counts <- read.csv(paste0(folder_4ce_files, "/non_psych_admit_inpt_11_17.csv"))
admit_all_codes <- read.csv(paste0(folder_4ce_files, "/psych_admit_all_codes.csv"))
admit_demog <- read.csv(paste0(folder_4ce_files, "/psych_admit_demog.csv"))
admit_inpt_primary <- read.csv(paste0(folder_4ce_files, "/psych_admit_inpt_primary.csv"))
```

Raw Data Counts

```{r}
nrow(unique(data.frame(admit_all_codes$patient_id)))
nrow(unique(data.frame(admit_demog$patient_id)))
```
### Read the ICD10 codes file

The file containing the ICD10 psychiatric related codes is located in
the `public-data` folder of the GitHub repository.

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   dplyr::select( disorder_group, ICD10_Code, description )
```

Old codes that Nelson used. These were updated to the ICD code list above in order to remove codes for accidental poisoning

```{r}
icdCodes_old <- read.csv("public-data/psychiatric_ICD-10_codes_adjusted.csv", header = TRUE, colClasses = "character")

colnames(icdCodes_old) <- c("disorder_group", "ICD10_Code", "description")

icdCodes_old <- icdCodes_old %>% 
  mutate(ICD10_Code = gsub("ICD10:", "", ICD10_Code))

excluded_codes <- icdCodes_old %>% anti_join(icdCodes)
qc_check <- admit_all_codes %>% filter(icd_10_code %in% excluded_codes$ICD10_Code) 
nrow(unique(data.frame(qc_check$patient_id)))
```

## Data-management

**Meghan Note: (Skip-to Lurie Data Formatting section)**

### Filtering on "inpatient" status

```{r eval=FALSE, include=FALSE}
# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
clinical_raw <- filter(clinical_raw, in_hospital == 1)
```

### Calculating duration of individual hospitalizations

Adding columns with hospitalization number and hospitalization length to the clinical raw file

```{r eval=FALSE, include=FALSE}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)

clinical_raw <- left_join(clinical_raw,
                          hospitalisations_seq_df,
                          by = c("patient_num", "cohort", "days_since_admission"))
```

# Lurie Data Formatting

**Clean up Lurie encounters**

As noted by Nelson: "some patients will have two encounters back-to-back. These should be considered the SAME encounter. The reason for having two encounters recorded is because medical admissions and psych admissions are considered separate admissions in our hospital. A simple way to deal with this redundancy is to check encounters for the same patient that start on the same day as the end of the prior encounter and merge those two into one encounter."

**First we will combine all of the psych + non-psych data pulls**

```{r}
# admit_demog$pysch = 1
# non_psych_inpt_counts$pysch = 0

admit_demog <- rbind(admit_demog, non_psych_inpt_counts)

## appears to be no overlap?
#View(admit_demog %>% distinct(patient_id, encounter_id, pysch) %>% group_by(psych) %>%  mutate(count = n()))
```

```{r}
## format dates
admit_demog$i_hsp_adm_dttm <- strptime(admit_demog$hsp_adm_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()
admit_demog$i_hsp_dsch_dttm <- strptime(admit_demog$hsp_dsch_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()

# # filter ages
# admit_demog <- admit_demog %>% filter(age_years > 10 & age_years < 18)  %>%
#  filter(i_hsp_adm_dttm >= start_date_plots & i_hsp_adm_dttm <= end_date_plots)

# unique encounter numbers should equal the row number (TRUE)
nrow(unique(data.frame(admit_demog$encounter_id))) == nrow(admit_demog)
nrow(admit_demog %>% distinct(patient_id, encounter_id))

encounters <- admit_demog %>% 
  mutate(gender = if_else(gender == "Male", "male", "female")) %>% 
  group_by(patient_id) %>% 
  arrange(patient_id, i_hsp_adm_dttm) %>% 
  filter(i_hsp_adm_dttm >= start_date_plots & i_hsp_adm_dttm <= end_date_plots) %>% 
  mutate(encounter_group = seq_along(encounter_id),
         day_lag = as.numeric(round(difftime(i_hsp_adm_dttm, lag(i_hsp_dsch_dttm), unit = "hours"), 2))) %>% 
  data.frame() %>%
  # adding 100 to temporarily handle the NAs
  mutate(day_lag = ifelse(is.na(day_lag), 100, day_lag),
  # here we want to ensure that if subsequent encounters occurred within a day (24 hours), that we impute the initial encounter_group as per Nelson's suggestion
         i_encounter_group = ifelse(day_lag <= 24, lag(encounter_group), encounter_group),
  consolidated = ifelse(day_lag <= 24, 1, 0))

# this is suggesting that 408 need to get consolidated?
table(encounters$consolidated)

check_lag <- encounters %>% filter(!day_lag == 100)
summary(check_lag$day_lag)
less24 <- check_lag %>% filter(day_lag <= 24) #449
summary(less24$day_lag)
hist(less24$day_lag)

encounters <- encounters %>%
  # in some cases there are more than two back-to-back encounters
  # to filter for these, first we will group by the patient_id and the new encounter group
  group_by(patient_id, i_encounter_group) %>%
  # next, we will calculate the latest admission date of this encounter
  mutate(max_date = max(i_hsp_adm_dttm)) %>%
  ungroup() %>%
  # next, we group just by Patient, and we identify the latest max_date
  # we then calculate the current time from the lag(max_date) and we will identify those within 24 hours as another encounter that should be grouped
  group_by(patient_id) %>%
  mutate(lag_date = lag(max_date),
         time_diff = as.numeric(round(difftime(i_hsp_adm_dttm, lag_date, unit = "hours"), 2))) %>%
  # adding 100 to temporarily handle the NAs
  mutate(time_diff = ifelse(is.na(time_diff), 100, time_diff),
    i_encounter_group = ifelse(time_diff <= 24, lag(i_encounter_group), i_encounter_group),
    consolidated2 = ifelse(time_diff <= 24, 1, 0),
  # recalculate the encounter sequence
  encounter_num = dense_rank(i_encounter_group)) %>% 
  group_by(patient_id, encounter_num)

# this is suggesting that 408 need to get consolidated?
table(encounters$consolidated2)

check_lag <- encounters %>% filter(!time_diff == 100)
summary(check_lag$time_diff)
less24 <- check_lag %>% filter(time_diff <= 24) #397 patients/admissions said to need to be consolidated
summary(less24$time_diff)
hist(less24$time_diff)

encounters <- encounters %>% 
  slice(1L)

nrow(encounters %>% distinct(patient_id, encounter_num))
```

Check to see if we should have combined any additional encounters

```{r}
test = encounters %>% 
       group_by(patient_id, encounter_num) %>% 
       mutate(max_date = max(i_hsp_adm_dttm)) %>% 
  ungroup() %>%
  group_by(patient_id) %>% 
  mutate(lag_date = lag(max_date),
         time_diff = as.numeric(difftime(i_hsp_adm_dttm, lag_date, unit = "hours")),
         check = if_else(time_diff <= 24 & encounter_num != lag(encounter_num), 1, 0))

#View(test %>% filter(check == 1))
```

**Add number of hospitalizations and LOS of stay for each patient**

```{r}
clinical_raw <- encounters %>% 
  group_by(patient_id) %>% 
  mutate(n_hospitalisation = n_distinct(as.character(encounter_num))) %>% 
  group_by(patient_id, encounter_num) %>% 
  # calculate the hospitalization length, while taking care to select the max(i_hsp_dsch_dttm) in the case that we have more than 1 row per the same encounter
  mutate(len_hospitalisation = round(as.numeric(difftime(max(i_hsp_dsch_dttm), min(i_hsp_adm_dttm), unit = "days"))),
         # replace last discharge with the true discharge datetime
         i_hsp_dsch_dttm = max(i_hsp_dsch_dttm),
         # replace age_years with earliest age on admission 
         age_years = min(age_years)) %>% 
  rename("date" = i_hsp_adm_dttm,
         "patient_num" = patient_id) %>% 
  group_by(patient_num, encounter_num) %>% 
  slice(1L) %>% 
  ungroup() %>% 
  select(patient_num, date, i_hsp_dsch_dttm, race, ethnicity, gender, age_years, encounter_num, n_hospitalisation, len_hospitalisation) %>% 
  distinct()

nrow(clinical_raw %>% distinct(patient_num, encounter_num))
```

### Counting ICD10 mental health codes in the obs_raw table

```{r echo=TRUE}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- admit_all_codes %>% dplyr::filter(icd_10_code == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", icd_10_code))
print(codesToReview)

### Calculate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( admit_all_codes %>%
  dplyr::filter( #concept_type == "DIAG-ICD10", 
                 icd_10_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( icd_10_code = as.character( icd_10_code ), 
                 concept_length = nchar( icd_10_code ), 
                 concept_check = ifelse( icd_10_code == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( icd_10_code, concept_length, concept_check) )

summary(as.factor( mentalHealthCodes$concept_check ) )

print( paste0("There are ", length(mentalHealthCodes$icd_10_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Calculating age at hospitalization time

```{r eval=FALSE, include=FALSE}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))
stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
```

### Applying inclusion criteria to filter and format the data

We are selecting hospitalizations based on the following criteria:

- Associated with at least 1 ICD code

- With a duration of hospitalization spanning over at least two consecutive days

- Starting on or after 2019/01/01

```{undefined eval=FALSE, include=FALSE}
obs_filtered_age <- obs_raw_age %>% 
  filter(concept_type == 'DIAG-ICD10', 
         days_since_admission >= 0)

clinical_filtered <- clinical_raw %>%
  filter( len_hospitalisation >= min_hosp_days, 
          cohort  %in% c("AllAdm")) %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic", "during_pandemic"))

if( time_period == "weeks"){
  clinical_filtered$time_p <- clinical_filtered$weeks
}else if(time_period == "days"){
  clinical_filtered$time_p <- clinical_filtered$date
}else if(time_period == "months"){
  clinical_filtered$time_p <- clinical_filtered$month
}

```

### Further filtering Lurie data

```{r}
clinical_raw$date <- strptime(clinical_raw$date, format="%Y-%m-%d") %>% as.POSIXct()

# create list of calendar_dates
date_ids <- seq(as.Date("2019-01-01"), as.Date(end_date_plots), 
                by= "day") %>% 
  data.frame() %>% 
  rename('calendar_date' = '.')

date_ids <- date_ids %>% 
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","during_pandemic"))

clinical_filtered <- clinical_raw %>% 
  filter(len_hospitalisation >= min_hosp_days) %>% 
  left_join(., date_ids, by = "date") %>% 
   dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                  weeks = as.Date(cut( date, breaks = "week")),
                  month = as.Date(cut( date, breaks = "month")),
                  year = format( date, "%Y"), 
                  period = ifelse( date < pandemic_start_date,
                                   "before_pandemic","during_pandemic"))

# check breaks example from BCH
example_bch <- read.delim2("example.txt", header = TRUE, sep = "")
example_lurie <- date_ids %>% distinct(calendar_date, weeks, month, year, period)

example_bch$calendar_date <- as.Date(example_bch$calendar_date)
example_bch$weeks <- as.Date(example_bch$weeks)
example_bch$month <- as.Date(example_bch$month)
example_bch$year <- as.character(example_bch$year)
example_bch <- example_bch %>% filter(!calendar_date >= "2021-06-01")

diff <- example_bch %>% anti_join(example_lurie)
nrow(diff)
diff <- example_lurie %>% anti_join(example_bch)
nrow(diff)
identical(example_bch, example_lurie)
all.equal(example_bch, example_lurie)

if( time_period == "weeks"){
  clinical_filtered$time_p <- clinical_filtered$weeks
}else if(time_period == "days"){
  clinical_filtered$time_p <- clinical_filtered$date
}else if(time_period == "months"){
  clinical_filtered$time_p <- clinical_filtered$month
}
```

```{r eval=FALSE, include=FALSE}
# Steps necessary to filter rows:
  # - encounters without ICD
  # - ICD diagnosed outside of encounters
  # - patients from the patients table that don't have any encounters matching the criteria
clinical_table <- clinical_filtered %>% 
  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation) %>% 
  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
# Adding new columns
observation_table <- obs_filtered_age %>% 
  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
demographic_table <- demo_raw %>% 
  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
all_codes <- observation_table %>% 
   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
```

**Create Lurie's Input_data table**

```{r}
nrow(unique(data.frame(clinical_filtered$patient_num)))
nrow(clinical_filtered %>% distinct(patient_num, encounter_num))

input_data <- clinical_filtered %>% 
  mutate(age_years = floor(age_years)) %>% 
  filter(date >= start_date_plots & date <= end_date_plots,
         age_years > 10 & age_years < 18) %>% 
  group_by(patient_num, encounter_num) %>% 
  arrange(date) %>% 
  slice(1L) %>% 
  ungroup() %>% 
  select(patient_num, encounter_num, race, ethnicity, gender, age_years, date, i_hsp_dsch_dttm, weeks, month, year, period, time_p, n_hospitalisation, len_hospitalisation) %>% 
  rename("sex" = gender,
         "age" = age_years)

nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))
```

```{r eval=FALSE, include=FALSE}
qc_check <- admit_all_codes %>% 
   rename("patient_num" = patient_id, 
          "concept_code" = diagnosis_code) %>% 
  left_join(., icdCodes, by = c("concept_code" = "ICD10_Code")) %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>% 
  left_join(input_data, ., by = "patient_num") %>% 
  mutate(i_diagnosis_dttm = strptime(diagnosis_dttm, format="%m/%d/%Y") %>% as.POSIXct(),
         date = format(as.POSIXct(date, format = "%m/%d/%Y %H:%M"), format = "%Y-%m-%d"),
         i_hsp_dsch_dttm = format(as.POSIXct(i_hsp_dsch_dttm, format = "%m/%d/%Y %H:%M"), format = "%Y-%m-%d")) %>% 
  dplyr::filter( diagnosis_code_source == "ICD-10-CM", 
                 #in_hospital == 1,
                 i_diagnosis_dttm >= date & i_diagnosis_dttm <= i_hsp_dsch_dttm,
                 age > 10 & age < 18,
                 date >= start_date_plots & date <= end_date_plots,
                 ) %>%
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)

```

```{r eval=FALSE, include=FALSE}
stopifnot(unique(all_codes$cohort) == "AllAdm")
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation >= min_hosp_days))
stopifnot(nrow(all_codes) == nrow(observation_table))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
```

### Filtering on age 11-17

```{r eval=FALSE, include=FALSE}
# Filtering hospitalizations where patient age is in inclusion criteria
observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)

# Filtering observation tables which patient_num and days_since_admission are still in clinical table after filtering on age

clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
            by = c("patient_num", "days_since_admission")) %>% 
  distinct(patient_num, n_hospitalisation, age_time_diagnosis)

clinical_table_ado <- clinical_table %>% 
  inner_join(clinical_test,
             by = c("patient_num", "n_hospitalisation"))

# Filtering observation tables which patient_num are still in clinical table after filtering on age
demographic_table_ado <- demographic_table %>% 
  inner_join(distinct(clinical_table_ado, patient_num), by = "patient_num")

all_codes_ado <- all_codes %>% 
   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
```

```{r eval=FALSE, include=FALSE}
stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
stopifnot(unique(all_codes_ado$in_hospital) == 1)
stopifnot(unique(all_codes_ado$len_hospitalisation >= min_hosp_days))
stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
```

Are there any patients in the admit_demog df who do not have any icd codes listed?

Note: this looks fine
```{r}
admit_demog <- read.csv(paste0(folder_4ce_files, "/psych_admit_demog.csv"))

# list of all patients who should have had at least 1 psych diagnosis
psych_ids <- admit_demog %>% 
  distinct(patient_id)
nrow(psych_ids)

# evaluate whether there are patients in admit_all_codes file that were not listed in the admit_demog file
admit_demog_ids <- admit_all_codes %>% 
  distinct(patient_id) %>% 
  filter(!patient_id %in% psych_ids$patient_id)
nrow(admit_demog_ids)

# evaluate whether there are patients in the admit_demog file that do not have diagnosis codes
psych_id_check <- psych_ids %>% 
  distinct(patient_id) %>% 
  filter(!patient_id %in% admit_all_codes$patient_id)
nrow(psych_id_check)
```

### Add the psychiatric/mental health ICD code description and filter

```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
#site <- unique(as.character(demo_raw$siteid))
site <- "Lurie"

icdCategory <- read.delim("public-data/icd10Codes.txt",
                          header = FALSE,
                          colClasses = "character")
colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")

# all_codes_desc <- all_codes_ado %>% 
#     left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
#   mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
#                                "psy", 
#                                "others")) %>%
#   dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
#                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
#                  period, time_p, disorder_group, description, icd_code_category )

# begin processing the ICD codes for the patients
all_codes_desc <- admit_all_codes %>% 
   rename("patient_num" = patient_id, 
          "concept_code" = icd_10_code) %>% 
  left_join(., icdCodes, by = c("concept_code" = "ICD10_Code")) %>% 
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others"))

nrow(input_data)
nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))

# Join mental health codes to the patient_number
input_data <- input_data %>% 
  left_join(., all_codes_desc, by = "patient_num") %>% 
  mutate(i_diagnosis_dttm = strptime(diagnosis_dttm, format="%m/%d/%Y") %>% as.POSIXct(),
         date = strptime(date, format="%Y-%m-%d") %>% as.POSIXct(),
         i_hsp_dsch_dttm = strptime(i_hsp_dsch_dttm, format="%Y-%m-%d") %>% as.POSIXct()) %>% 
  dplyr::filter(date >= start_date_plots & date <= end_date_plots)

nrow(input_data)
nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))

# we will temporarily split the data frame between our psych patients with ICD codes and our non-pysch without, due to the filtering function automatically dropping rows with missing ICD codes
# of note: these are just some patients who have an ICD code - could be some non-psych codes as the original query including some icd codes for poisoning that ended up being excluded
keep_psych <- input_data %>% 
  group_by(patient_num, encounter_num) %>% 
  filter(i_diagnosis_dttm >= date & i_diagnosis_dttm <= i_hsp_dsch_dttm) %>% 
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)

keep_non_psych = input_data %>% 
  group_by(patient_num, encounter_num) %>% 
  filter(is.na(i_diagnosis_dttm)) %>% 
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)

# keep non_psych whose dx date did not occur during the admission
# but correct the diagnosis data to match the non_psych group
keep_non_psych_v2 = input_data %>% 
  group_by(patient_num, encounter_num) %>% 
  filter(i_diagnosis_dttm < date | i_diagnosis_dttm >= i_hsp_dsch_dttm) %>% 
  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category) %>% 
  mutate(concept_code = NA,
         disorder_group = NA,
         description = NA, 
         icd_code_category = NA)

input_data <- rbind(keep_psych, keep_non_psych, keep_non_psych_v2)

nrow(input_data)
nrow(unique(data.frame(input_data$patient_num)))
nrow(input_data %>% distinct(patient_num, encounter_num))

all_codes_desc <- input_data %>% 
  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))

all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others"),
          icd_code_category = ifelse(is.na(icd_code_category), "others", icd_code_category))

nrow(all_codes_desc %>% distinct(patient_num, encounter_num))

all_codes_psy <- all_codes_desc %>% 
  filter(concept_code %in% icdCodes$ICD10_Code)

nrow(unique(data.frame(all_codes_psy$patient_num)))
nrow(all_codes_psy %>% distinct(patient_num, encounter_num))
# View(all_codes_psy %>% 
#        ungroup() %>% 
#        distinct(patient_num, encounter_num, time_p, period) %>% 
#        group_by(period) %>% 
#        mutate(count = n_distinct(patient_num, encounter_num)) %>% 
#        distinct(period, count))

## how many weeks in 'during' vs 'before' pandemic (15 for both)
# View(all_codes_psy %>% 
#        ungroup() %>% 
#        distinct(month, period) %>%
#        group_by(period) %>% 
#        mutate(count = n_distinct(month)))

#all_codes_desc$i_date <- strptime(all_codes_desc$date, format="%Y-%m-%d") %>% as.POSIXct()

# if( raceAvailable == TRUE ){
#   all_codes_psy <- all_codes_psy %>%
#     left_join( race_raw, by = c("patient_num") )
# }
```

Evaluate numbers before vs after pandemic

```{r}
before_v_after <- all_codes_desc %>% 
  distinct(patient_num, encounter_num, period, icd_code_category) %>% 
  group_by(period, icd_code_category) %>% 
  mutate(count = n_distinct(patient_num, encounter_num)) %>% 
  distinct(period, icd_code_category, count)

before_v_after
```

QC Checks for Psych patients

```{r}
p1 <- ggplot(all_codes_desc %>%
               filter(icd_code_category == "psy") %>% 
               group_by(weeks, period) %>% 
               arrange(weeks) %>% 
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      weeks = as.Date(weeks)) %>% 
               distinct(weeks, psych_any_count, period),
       aes(x = weeks, y = psych_any_count, fill = period)) + 
  geom_bar(stat="identity", binwidth=1) + 
  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
  scale_color_brewer(palette="Dark2", direction = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1) + 
  ylab("Any Psychiatric Disease") + 
  xlab("") + 
  #scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

p1

# add an extra day to distinguish the before and during which occur in same month

p1 <- ggplot(all_codes_desc %>% 
               filter(icd_code_category == "psy") %>% 
               group_by(month) %>% 
               arrange(month) %>% 
               mutate(month = as.character(month), 
                 month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month)) %>% 
               group_by(month, period) %>% 
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      month = as.Date(month)) %>%
               ungroup() %>% 
               distinct(month, psych_any_count, period),
       aes(x = month, y = psych_any_count, fill = period)) + 
  geom_bar(stat="identity", width = 25) + 
  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
  scale_color_brewer(palette="Dark2", direction = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1) + 
  ylab("Any Psychiatric Disease") + 
  xlab("") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

p1

# p1 <- ggplot(all_codes_desc %>% 
#                filter(icd_code_category == "psy") %>% 
#                group_by(month) %>% 
#                arrange(month) %>% 
#                mutate(month = as.character(month), 
#                  period = ifelse(month == "2020-03-01", "March", period)) %>% 
#                group_by(month, period) %>% 
#                mutate(psych_any_count = n_distinct(patient_num, encounter_num),
#                       month = as.Date(month)) %>%
#                ungroup() %>% 
#                distinct(month, psych_any_count, period),
#        aes(x = month, y = psych_any_count, fill = period)) + 
#   geom_bar(stat="identity", width = 25) + 
#   geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
#   scale_color_brewer(palette="Dark2", direction = -1) +
#   scale_fill_brewer(palette = "Dark2", direction = -1) + 
#   ylab("Any Psychiatric Disease") + 
#   xlab("") + 
#   scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
#   theme_bw() + 
#   theme(legend.title = element_blank(),
#         axis.text.x = element_text(angle = 45, hjust = 1))
# 
# p1

# remove march
p1 <- ggplot(all_codes_desc %>% 
               filter(icd_code_category == "psy") %>% 
               group_by(month) %>% 
               arrange(month) %>% 
               mutate(month = as.character(month)) %>% 
               filter(!month == "2020-03-01") %>% 
                 #month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month)) %>% 
               group_by(month, period) %>% 
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      month = as.Date(month)) %>%
               ungroup() %>% 
               distinct(month, psych_any_count, period),
       aes(x = month, y = psych_any_count, fill = period)) + 
  geom_bar(stat="identity", width = 25) + 
  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
  scale_color_brewer(palette="Dark2", direction = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1) + 
  ylab("Any Psychiatric Disease") + 
  xlab("") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

p1
```

### Psych vs Non-psych

```{r}
p1 <- ggplot(all_codes_desc %>%
               #filter(icd_code_category == "psy") %>% 
               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
               # group_by(month) %>% 
               # mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
               # group_by(patient_num, month) %>% 
               # filter(check == max(check)) %>% 
               # slice(1) %>% 
               group_by(month, period, icd_code_category) %>% 
               arrange(month) %>% 
               mutate(month = as.character(month)) %>%
               filter(!month == "2020-03-01") %>%
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      month = as.character(month), 
                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
                      month = as.Date(month)) %>% 
               distinct(month, icd_code_category, psych_any_count, period) %>% 
               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
               mutate(sum = psy + others,
                      perc_psyc = psy/sum),
       aes(x = month, y = perc_psyc, fill = period)) + 
  geom_bar(stat="identity", width = 20) + 
  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
  scale_color_brewer(palette="Dark2", direction = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1) + 
  ylab("Any Psychiatric Disease") + 
  xlab("") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
  #facet_wrap(~icd_code_category) +
  theme_bw() + 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

p1

p1 <- ggplot(all_codes_desc %>%
               #filter(icd_code_category == "psy") %>% 
               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
               # group_by(month) %>% 
               # mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
               # group_by(patient_num, month) %>% 
               # filter(check == max(check)) %>% 
               # slice(1) %>% 
               group_by(month, period, icd_code_category) %>% 
               arrange(month) %>% 
               # mutate(month = as.character(month)) %>% 
               # filter(!month == "2020-03-01") %>% 
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      month = as.character(month), 
                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
                      month = as.Date(month)) %>% 
               distinct(month, icd_code_category, psych_any_count, period) %>% 
               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
               mutate(sum = psy + others,
                      perc_psyc = psy/sum),
       aes(x = month, y = perc_psyc, fill = period)) + 
  geom_bar(stat="identity", width = 20) + 
  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
  scale_color_brewer(palette="Dark2", direction = -1) +
  scale_fill_brewer(palette = "Dark2", direction = -1) + 
  ylab("Any Psychiatric Disease") + 
  xlab("") + 
  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
  #facet_wrap(~icd_code_category) +
  theme_bw() + 
  theme(legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))

p1

test <- all_codes_desc %>%
               #filter(icd_code_category == "psy") %>% 
               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
               # group_by(month) %>% 
               # mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
               # group_by(patient_num, month) %>% 
               # filter(check == max(check)) %>% 
               # slice(1) %>% 
               group_by(month, period, icd_code_category) %>% 
               arrange(month) %>% 
               mutate(psych_any_count = n(),
                      month = as.character(month), 
                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
                      month = as.Date(month)) %>% 
               distinct(month, icd_code_category, psych_any_count, period) %>% 
               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
               mutate(sum = psy + others,
                      perc_psyc = psy/sum)

before <- test %>% filter(period == "before_pandemic")
during <- test %>% filter(period == "during_pandemic")

summary(before$perc_psyc)
summary(during$perc_psyc)

# don't divide by month

#unique patient and encounter by month
test <- all_codes_desc %>%
               #filter(icd_code_category == "psy") %>% 
               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
               group_by(month) %>% 
               mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
               group_by(patient_num, month) %>% 
               filter(check == max(check)) %>% 
               slice(1) %>% 
               group_by(period, icd_code_category) %>% 
               arrange(month) %>% 
               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
                      month = as.character(month), 
                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
                      month = as.Date(month)) %>% 
               distinct(icd_code_category, psych_any_count, period) %>% 
               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
               mutate(sum = psy + others,
                      perc_psyc = psy/sum)

before <- test %>% filter(period == "before_pandemic")
during <- test %>% filter(period == "during_pandemic")

summary(before$perc_psyc)
summary(during$perc_psyc)

```

### Add the non-pysch patient encounters

```{r eval=FALSE, include=FALSE}
non_psych_inpt_counts$date <- strptime(non_psych_inpt_counts$hsp_adm_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()

# filter age and study inclusion period
non_psych_inpt_counts <- non_psych_inpt_counts %>% 
  filter(len_hospitalisation >= min_hosp_days) %>% 
  filter(age_years > 10 & age_years < 18,
         date >= start_date_plots & date <= end_date_plots) %>% 
  dplyr::mutate(#date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic", "during_pandemic"))

if( time_period == "weeks"){
  non_psych_inpt_counts$time_p <- non_psych_inpt_counts$weeks
}else if(time_period == "days"){
  non_psych_inpt_counts$time_p <- non_psych_inpt_counts$date
}else if(time_period == "months"){
  non_psych_inpt_counts$time_p <- non_psych_inpt_counts$month
}

# we will then indicate that these patients do not have psychiatric diseases
non_psych_inpt_counts <- non_psych_inpt_counts %>% 
  rename(age = "age_years") %>%  
  mutate(patient_num = row_number() + 78332,
         encounter_num = NA, 
         sex = NA,
         concept_code = NA, 
         n_hospitalisation = NA, 
         len_hospitalisation = NA, 
         disorder_group = NA,
         description = NA,
         icd_code_category = "others", 
         icd_code_letter = NA,
         icd_description_category = NA) %>% 
  select(colnames(all_codes_desc))
```

Combine the psych and non-psych dataframes

```{r eval=FALSE, include=FALSE}
all_codes_desc <- rbind(all_codes_desc, non_psych_inpt_counts)
```

#### QC step: do a summary of the mental health ICD codes present in each site
As output we will generate a table with: the number of observations per code, number of patients diagnosed with each code and the date range in which the code was used.

```{r}
#filter the whole dataset to only mental codes
mental_codes_qc <- all_codes_desc %>% 
  filter( concept_code %in% mentalHealthCodes$icd_10_code)

mental_codes_qc_output <- mental_codes_qc %>%
  group_by( concept_code ) %>%
  dplyr::summarise( distinct_patients = ifelse(length(unique(patient_num)) > obfuscation | isFALSE(obfuscation),
                               length(unique(patient_num)), 
                               0.5),
                    distinct_observations = ifelse(length(patient_num) > obfuscation | isFALSE(obfuscation),
                               length(patient_num), 
                               0.5),                           
                    min_date = min( date ), 
                    max_date = max( date ))%>% 
  arrange(desc(distinct_observations))
mental_codes_qc_output <- left_join( mental_codes_qc_output, icdCodes, by=c("concept_code"= "ICD10_Code")) 
mental_codes_qc_output  <- mental_codes_qc_output[, c("disorder_group", "description", "concept_code", "distinct_patients", 
                                                    "distinct_observations", "min_date", "max_date")]
mental_codes_qc_output[1:10, ]
```

# Results

## Admission

First calculation: total number of patients, matching the inclusion criteria, admitted on one day (without filtering by any disease subtype). Summed up over one week, or one month depending on the time unit chosen. Not filtered on age.

```{r}
admission_plot_count <- function(df,
                               x,
                               y, 
                               title,
                               y_title,
                               size = NULL, 
                               color = NULL){
  plot <- ggplot(df, aes_string(x = x, y = y, color = color, size = size))  +
    geom_point() +
    scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y", limits = c(start_date_plots, end_date_plots), expand = expansion(mult = 0.01, add = 1)) +
    theme(axis.text.x=element_text(angle=60, hjust=1))+
    labs(x = "Calendar date",
         y = y_title, 
       title = title) 
  return(plot)
}

```

```{r}
new_hospitalizations <- all_codes_desc %>%
  group_by(patient_num, n_hospitalisation) %>%
  slice(which.min(time_p)) %>%
  #filter( in_hospital == 1 ) %>%
    group_by( time_p ) %>%
  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>% 
  ungroup()

nh <- admission_plot_count(df = new_hospitalizations, 
                     x = "time_p",
                     y = "count", 
                     title = paste0("New admissions (per ", time_period,")"), 
                     y_title = "Patient Count")

nh
```

## New admissions age range 11-17

```{r}
new_hospitalizations_ado <- all_codes_desc %>%
  group_by(patient_num, n_hospitalisation) %>%
  slice(which.min(time_p)) %>%
  #filter( in_hospital == 1 ) %>%
    group_by( time_p ) %>%
  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>% 
  ungroup()

nh_ado <- admission_plot_count(df = new_hospitalizations_ado, 
                     x = "time_p",
                     y = "count", 
                     title = paste0("New admissions (per ", time_period, ")"), 
                     y_title = "Patient Count")+
    labs( subtitle = "(age range 11-17)")

nh_ado
```

## Total hospitalize patients age range 11-17

Calculating the total number of patients hospitalized on a given day (taking into account the hospitalization length).

```{r}
unique_patients_hospitalized <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

th <- admission_plot_count(df = unique_patients_hospitalized, 
                     x = "time_p",
                     y = "count", 
                     title = paste0("Total hospitalization (per ", time_period, ")"), 
                     y_title = "Patient Count") +
  labs( subtitle = "(age range 11-17)")
th
```

## Patient count filtered on psychiatric disorders

### New admissions

```{r}
count_icd_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p) %>%
  summarise(count_psy =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

count_icd <- count_icd %>%
  left_join(count_icd_psy, 
            by = c("time_p", "period")) %>% 
  replace_na(list(count_psy = 0))

count_icd$percentage_psy <- 100*count_icd$count_psy / count_icd$count_icd

thpd_psyCount <- count_icd %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
  geom_point() +
  # geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

thpd_psyCount

```

### Ratio: patients with psychiatric condition / patients without psychiatric condition

```{r}
all_patients <- unique( all_codes_desc[, c("patient_num", "time_p")])
nrow(unique(data.frame(all_patients$patient_num)))
patients_with_psy_conditions <- unique( all_codes_psy[, c("patient_num", "time_p")] )
nrow(unique(data.frame(patients_with_psy_conditions$patient_num)))
patients_without_psy_conditions <- setdiff(all_patients,patients_with_psy_conditions)
nrow(unique(data.frame(patients_without_psy_conditions$patient_num)))

patients_without_psy_conditions_counts <- patients_without_psy_conditions %>%
  group_by( time_p ) %>%
  summarise(count_no_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic"))

patients_with_psy_conditions_counts <-  patients_with_psy_conditions %>%
  group_by( time_p ) %>%
  summarise(count_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic"))

  ratios_patients_with_without_psy <- patients_without_psy_conditions_counts %>%
  full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
  mutate( ratio = count_psy_patients/count_no_psy_patients, 
          percentage = 100 * count_psy_patients/(count_no_psy_patients + count_psy_patients))

```

```{r}
ratios_patients_with_without_psy %>%
  pivot_longer(cols = c("count_no_psy_patients", "count_psy_patients"), 
               values_to = "count", 
               names_to = "count_type") %>% 
  ggplot(aes(x = time_p, y = count, color = count_type)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Count",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Count comparison: patients with vs without psychiatric conditions (per ", time_period,")")) +
  guides(color = guide_legend(title="Patient diagnoses types"))
```

```{r}
ratio_psy_non_psy <- ratios_patients_with_without_psy %>%
  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Ratio (with psy conditions/without psy conditions)",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))

ratio_psy_non_psy

```

```{r}
ratios_patients_with_without_psy %>%
  ggplot(aes(x = time_p, y = percentage, color = period)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
  guides(color = guide_legend(title="Patient diagnoses types"))
```

## Interrupted Time-Series Analyses

```{r}
# Model without seasonality
time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
  output_model <- broom::tidy(model)
  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
  pred_model <- predict(model, df_count_psy, type = "response", interval = "confidence") %>% as.data.frame()
  pred_model_b <- predict(model, mutate(df_count_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
    as.data.frame()
  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
  if (clearance_period) {
    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
                                          "before_pandemic",
                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
                                                 "during_pandemic",
                                                 "during_clearance")
    )
    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
                                             pred_model$fit_wo_intervention,
                                             NA)
  } else {
    pred_model$clearance_period <- pred_model$period
  }
  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
    geom_point(aes(colour = clearance_period)) +
    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
    scale_color_manual(name = "Time Series projection", values = c("before_pandemic" = "black",
                                                                   "Expected mean w/o pandemic" = "red",
                                                                   "during_pandemic" = "black",
                                                                   "during_clearance" = "white")) +
    annotate("text", x = as.Date("2019-10-01"), y = max(pred_model[[count_col]]) + 5, label = annotation) +
    theme(legend.position = "bottom")

  if (clearance_period) {
    plot +
      geom_vline(xintercept = as.Date(max(pred_model$time_p[pred_model$clearance_period == "before_pandemic"])), linetype = "dashed") +
      geom_vline(xintercept = as.Date(min(pred_model$time_p[pred_model$clearance_period == "during_pandemic"])), linetype = "dashed")
  } else {
    plot + geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")
  }
}
```

```{r}
# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
patient_count_psy_period_psy <- ratios_patients_with_without_psy %>%
  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
  mutate(time = seq_along(time_p),
         month_year = format(time_p, "%m")
  )

start_clear_period <- as.Date("2020-03-15")
end_clear_period <- as.Date("2020-05-15")
patient_count_psy_period_psy_clear <- patient_count_psy_period_psy %>%
  mutate( count_no_psy_patients = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                             count_no_psy_patients,
                             NA), 
          count_no_psy_patients = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                             count_no_psy_patients,
                             NA), 
          ratio = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                           time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                         ratio,
                         NA),
          percentage = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                           time_p > as.Date(cut( end_clear_period, breaks = time_period)),
                         percentage,
                         NA)
  )
```

```{r, its count based}
ts_linear_count <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy)
ts_linear_count_clear <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy_clear)
ts_seasonal_count <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
ts_seasonal_count_clear <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
```

```{r, its percentage based}
ts_linear_percent <- lm(percentage ~ period * time, patient_count_psy_period_psy)
ts_linear_percent_clear <- lm(percentage ~ period * time, patient_count_psy_period_psy_clear)
ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
ts_seasonal_percent_clear <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
```

### ITS: patient counts, no clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_count,
                              patient_count_psy_period_psy,
                              "count_psy_patients",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: patient counts, clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_count_clear,
                              patient_count_psy_period_psy,
                              "count_psy_patients",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: patient counts, no clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_count,
                              patient_count_psy_period_psy,
                              "count_psy_patients",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: patient counts, with clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_count_clear,
                              patient_count_psy_period_psy,
                              "count_psy_patients",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Count")
```

### ITS: psy percentage, no clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_percent,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS: psy percentage, clearance period, no seasonality effect

```{r}
time_series_analysis_plotting(ts_linear_percent_clear,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS: psy percentage, no clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_percent,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "Without seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS: psy percentage, with clearance period, with seasonality effect

```{r}
time_series_analysis_plotting(ts_seasonal_percent_clear,
                              patient_count_psy_period_psy,
                              "percentage",
                              clearance_period = TRUE) +
  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
       subtitle = "With seasonality effect",
       x = time_period,
       y = "Percentage")
```

### ITS bootstrap

#### Parametric bootstrap

```{r}
bootstrap_size <- 1000
binom_proportion <- function(n, prob, size) {
  (rbinom(n, size, prob) / size) * 100
}
bootstrapped_timepoints <- mapply(binom_proportion,
                                  n = bootstrap_size,
                                  size = patient_count_psy_period_psy$count_no_psy_patients +
                                    patient_count_psy_period_psy$count_psy_patients,
                                  prob = patient_count_psy_period_psy$percentage/100)

bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
for (iteration in 1:bootstrap_size) {
  df_iteration <- data.frame(period = patient_count_psy_period_psy$period,
                             time = patient_count_psy_period_psy$time,
                             percentage = bootstrapped_timepoints[iteration, ])
  ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
  linear_model <- broom::tidy(ts_linear_percent) %>%
    filter(term %in% c("time", "periodduring_pandemic:time"))
  coefficients <- linear_model$estimate
  names(coefficients) <- linear_model$term
  bootstrapped_coefficients_list[[iteration]] <- coefficients
  df_iteration$predicted <- predict(ts_linear_percent)
  df_iteration$time_p <- patient_count_psy_period_psy$time_p
  bootstrapped_sample_list[[iteration]] <- df_iteration
}

bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration")

quantile_bootstrapped_fitted <- bootstrapped_fitted_df %>%
  group_by(time_p, period) %>%
  summarise("2.5%" = quantile(predicted, 0.025),
            "50%" = quantile(predicted, 0.5),
            "97.5%" = quantile(predicted, 0.975)) %>%
  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
               names_to = "quantiles")

bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
mean(bootstrapped_coefficients_df$time)
mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
# pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
#                                    "< 1e-4",
#                                    str(pvalue_coeff_interaction))
if (pvalue_coeff_interaction == 0) {
  pvalue_coeff_interaction <- 1e-4
  print("equals 0")
} else {
  pvalue_coeff_interaction <- as.character(pvalue_coeff_interaction)
}

plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
                     round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
                     " (pvalue: ",
                     pvalue_coeff_interaction,
                     ")")

ggplot(bootstrapped_fitted_df, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
  geom_line(aes(color = period),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted,
            aes(y = value, group = interaction(quantiles, period)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted[quantile_bootstrapped_fitted$quantiles == "50%", ],
            aes(y = value, group = period)) +
  geom_point(data = patient_count_psy_period_psy,
             aes(y = percentage),
             alpha = 0.5) +
  scale_colour_manual(values = cbPalette) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  annotate("text",
           x = as.Date("2019-10-01"),
           y = max(patient_count_psy_period_psy[["percentage"]]) + 5,
           label = plot_annotation) +
  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period")

```

## Psychiatric conditions aggregating by subcategories

### New admissions: count

```{r}
nrow(unique(data.frame(all_codes_psy$patient_num)))
count_disorder_group <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p, disorder_group ) %>%
  summarise(count_dg =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

dg <- count_disorder_group %>%  
  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient Count",
       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg
```

#### Subset: the most prevalent disorder group
```{r}
dg_subset <- count_disorder_group %>%
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient Count",
       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 12),
        axis.text.x = element_text(size = 12, angle = 90), 
        axis.text.y = element_text(size = 11), 
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 11),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dg_subset
```

### New admissions: percentage

```{r}
# estimate the counts
count_icd_all <- all_codes_desc %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5)) %>%
  mutate( period = ifelse( time_p < pandemic_start_date,
                           "before_pandemic",
                           "during_pandemic")
  )

# join the total with the disorder group count to estimate the percentage
perc_disorder_group <- count_icd_all %>%
  left_join(count_disorder_group, 
            by = c("time_p", "period")) %>% 
  replace_na(list(count_dg = 0))

perc_disorder_group$percentage_dg <- 100*perc_disorder_group$count_dg / perc_disorder_group$count_icd

dgPerc <- perc_disorder_group %>%  
  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient (%)",
       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc
```

```{r}
dgPerc_subset <- perc_disorder_group %>%  
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient (%)",
       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 12),
        axis.text.x = element_text(size = 12, angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])

dgPerc_subset
```

## Psychiatric conditions, dichotomized on patients' sex

### Count

```{r}
# counts per time period and sex
patients_hospitalized_agg_sex <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( time_p, sex ) %>%
  summarise(count_sex =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

patients_hospitalized_agg_sex %>%
  ggplot(aes(x = time_p, y = count_sex, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = time_period,
       title = paste0("Patient counts per ", time_period, " aggregated by sex"),
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

### Percentage

```{r}
##percentage
total_patients_psy <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots ) %>%
    group_by( time_p ) %>%
  summarise(count =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

patients_hospitalized_agg_sex_perc <- total_patients_psy %>%
  left_join( patients_hospitalized_agg_sex, by = c("time_p")) %>% 
  replace_na(list(count_sex = 0))

patients_hospitalized_agg_sex_perc$percentage <- 100*patients_hospitalized_agg_sex_perc$count_sex/patients_hospitalized_agg_sex_perc$count

patients_hospitalized_agg_sex_perc %>%
  ggplot(aes(x = time_p, y = percentage, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Percentage ",
       x = time_period,
       title = paste0("Patient (%) per ", time_period," aggregated by sex"),
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

## Table 1

```{r}

t.test2 <- function(m1, m2, s1, s2, n1, n2) {
  # t.test computed from sample statistics, ie using mean, standard deviation and count value of the samples
  # useful in case we can only retrieve aggregated stats from the sites
  # ref: https://stats.stackexchange.com/questions/30394/how-to-perform-two-sample-t-tests-in-r-by-inputting-sample-statistics-rather-tha
  # and https://en.wikipedia.org/wiki/Student%27s_t-test?oldformat=true#Equal_or_unequal_sample_sizes,_unequal_variances_(sX1_%3E_2sX2_or_sX2_%3E_2sX1)
  se <- sqrt( (s1^2/n1) + (s2^2/n2) )
  # welch-satterthwaite df
  df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
  t <- (m1-m2)/se
  dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))
  names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
  return(dat)
}

nrow(unique(data.frame(all_codes_psy$patient_num)))

age_by_period <- all_codes_psy %>%
  distinct(patient_num, age, period) %>% 
  dplyr::group_by(period) %>%
  dplyr::summarise( mean = mean( age ),
                    sd = sd( age ),
                    count = n_distinct(patient_num))

age_test <- t.test2(m1 = age_by_period[[1, "mean"]],
                    m2 = age_by_period[[2, "mean"]],
                    s1 = age_by_period[[1, "sd"]],
                    s2 = age_by_period[[2, "sd"]],
                    n1 = age_by_period[[1, "count"]],
                    n2 = age_by_period[[2, "count"]]) %>%
  `[[`("p-value")
```

```{r}
sex <- all_codes_psy %>%
  dplyr::distinct(patient_num, sex) %>% 
  dplyr::group_by( sex ) %>%
  dplyr::summarise( count = n_distinct(patient_num) ) %>%
  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
                 total = sum( count ), percentage = round( count/total*100, 2)  )

sex_by_period <- all_codes_psy %>%
  dplyr::distinct(patient_num, period, sex) %>% 
  group_by(period, sex ) %>%
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, .add=TRUE) %>%
  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
          percentage=round(count/sum( count )*100, 2)) %>%
  replace_na(list(male = 0, female = 0))

sex_test <- sex_by_period %>%
  select(-percentage) %>%
  pivot_wider(names_from = sex, values_from = count) %>%
  ungroup() %>%
  select(male, female) %>%
  chisq.test() %>%
  `[[`("p.value")
```

```{r}
summary_diagnosis <- all_codes_psy %>%
  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>%
  dplyr::group_by( concept_code, description, period ) %>%
  dplyr::count( patient_num ) %>%
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc(count) ) %>%
  dplyr::ungroup()

summary_diagnosis_test <- summary_diagnosis %>%
  tidyr::pivot_wider(names_from = period,
                     values_from = count) %>%
  #select(late_pandemic, early_pandemic) %>%
  select(during_pandemic) %>%
  #tidyr::replace_na(list(late_pandemic = 0, early_pandemic = 0)) %>%
  tidyr::replace_na(list(during_pandemic = 0)) %>%
  chisq.test() %>%
  `[[`("p.value")

```

```{r}
summary_disorder_group <- all_codes_psy %>%
  dplyr::group_by( disorder_group ) %>%
  dplyr::count( patient_num ) %>%
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc( count ))

summary_disorder_group_by_period <- all_codes_psy %>%
  dplyr::distinct( patient_num, disorder_group, period, weeks ) %>%
  dplyr::group_by( period, disorder_group ) %>%
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, .add=TRUE) %>%
  mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5),
         percentage = count/sum( count )*100)

summary_disorder_group_by_period_test <- summary_disorder_group_by_period %>%
  pivot_wider(id_cols = disorder_group, names_from = period, values_from = c(count, percentage)) %>%
  mutate(across(where(is.numeric), ~replace_na(.x, replace = 0))) %>%
  #select(count_early_pandemic, count_late_pandemic) %>%
  select(count_during_pandemic) %>%
  chisq.test() %>%
  `[[`("p.value")

```

```{r}
format_counts <- function(count, percentage) {
  glue::glue("{count} ({percentage} %)", count = count, percentage = percentage)
}

age_table1 <- age_by_period %>% 
  mutate(result = glue::glue("{mean} (sd: {sd})", mean = round(mean, 2), sd = round(sd, 2))) %>% 
  select(result, period) %>% 
  pivot_wider( names_from = period, values_from = result ) %>%
  mutate(name = "Age")

sex_table1 <- sex_by_period %>%
  bind_rows(
    sex_by_period %>%
      group_by(period) %>%
      dplyr::summarise(sex = "Total", count = sum(count), percentage = sum(percentage))
  ) %>%
  mutate(result = format_counts(count, percentage)) %>%
  select(-c(count, percentage)) %>%
  pivot_wider( names_from = period, values_from = result ) %>%
  rename(name = sex)

disorder_table1 <- summary_disorder_group_by_period %>%
  arrange(desc(count)) %>% 
  bind_rows(
    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
  ) %>%
  mutate(percentage = round(percentage, 2)) %>% 
  mutate(result = format_counts(count, percentage)) %>%
  select(-c(count, percentage)) %>%
  pivot_wider( names_from = period, values_from = result ) %>%
  replace_na( list(during_pandemic = "0 (0 %)")) %>%
  rename(name = disorder_group)
```

```{r}
list_tests <- c("Sex" = sex_test,
                "Age" = age_test,
                "Diagnoses" = summary_diagnosis_test,
                "Disorder groups" = summary_disorder_group_by_period_test
)

list_table1 <- list(
  "Sex" = sex_table1,
  "Age" = age_table1,
  "Disorder groups" = disorder_table1)

tests <- list_tests %>%
  data.frame() %>%
  round(2)
names(tests) <- "pvalue"
tests["Category"] <- row.names(tests)

table1 <- list_table1 %>%
  lapply(function(df) mutate(df, across(.fns = as.character))) %>%
  dplyr::bind_rows(.id = "Category") %>%
  dplyr::left_join(tests, by = "Category")

table1 %>%
  group_by(Category) %>%
  mutate(pvalue = ifelse(duplicated(pvalue, fromLast = TRUE), "", as.character(pvalue))) %>%
  ungroup() %>%
  mutate(Category = "") %>%
  mutate(across(.cols = c(before_pandemic, during_pandemic), .fns = replace_na, "0")) %>%
  rename_with(str_to_title) %>%
  rename_with(sub, pattern = "_", replacement = " ") %>%
  mutate(across(.fns = str_to_title)) %>%
  mutate(across(.fns = sub, pattern = "_", replacement = " ")) %>%
  kable( booktabs = T) %>%
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
```

# Save the output in a RData file. 
```{r}
length_hospitalisation_values <- all_codes_psy %>% 
  dplyr::distinct( patient_num, encounter_num, n_hospitalisation, len_hospitalisation, period ) %>%
  dplyr::mutate( siteid = site ) %>%
  dplyr::select( len_hospitalisation, siteid, period )

# de-identify to share with alba/arnaud
to_share <- length_hospitalisation_values %>% 
  ungroup() %>% 
  mutate(siteid = "Lurie") %>% 
  select(len_hospitalisation, siteid, period)

write.csv(to_share, "../2022.04.06_Lurie_Output/Lurie_los.csv", row.names = FALSE)

# add the site id as column for each of the output data frames
count_icd$siteid = site 
patient_count_psy_period_psy$siteid = site 
patient_count_psy_period_psy_clear$siteid = site 
bootstrapped_coefficients_df$siteid = site 
bootstrapped_fitted_df$siteid = site 
perc_disorder_group$siteid = site 
patients_hospitalized_agg_sex_perc$siteid = site 
table1$siteid = site 
ratios_patients_with_without_psy$siteid = site 
mental_codes_qc_output$siteid = site

mental_codes_qc_output <- mental_codes_qc_output %>% select(-min_date, -max_date)

save( count_icd,
      mental_codes_qc_output,
      ratios_patients_with_without_psy,
      patient_count_psy_period_psy,
      patient_count_psy_period_psy_clear,
      bootstrapped_coefficients_df,
      bootstrapped_fitted_df,
      perc_disorder_group,
      patients_hospitalized_agg_sex_perc,
      table1,
      #length_hospitalisation_values,
      file = paste0("../2022.04.06_Lurie_Output/peds_psy_aggResults_", site, ".RData"))

rm(length_hospitalisation_values)
```